---
title: "CVD Risk Calculator"
output: html_notebook
---

*Author: Koki*  
*Date: Last edited on 3/21/2019*

The following notebook calculates the CVD Risk score as shown in the following paper:
> Conroy, R.M., Pyörälä, K., Fitzgerald, A.P., Sans, S., Menotti, A., De Backer, G., De Bacquer, D., Ducimetière, P., Jousilahti, P., Keil, U., et al. (2003). Estimation of ten-year risk of fatal cardiovascular disease in Europe: The SCORE project. European Heart Journal 24, 987–1003.

The paper describes the **SCORE project**, which was initiated as an effort to develop a risk scoring system for use in clinical management of cardiovascular risk in European clinical practice. The project was conducted on a pool of dataset comprising of 12 European cohort studies, a total of 205178 participant data. The 10 year risk of fatal cardiovascular disease was calculated using a **Weibull model**, where age was used as a measure of exposure time to a risk factor. 

The underlying formula for the score calculation is as follows: 

### 0. Load the Coefficient tables
**TO DO: Figure out how to set up column levels**

The coefficients come from the trained model fromt the SCORE project
```{r}
# Load coefficients for coronary heart disease
risk_table_A <- data.frame(risk_lvl = c('Low risk', 'Low risk', 'High risk', 'High risk'),
                           sex = c('Men', 'Women'),
                          # chd = c('CHD', 'Non-CHD CVD'),
                           a1 = c(-22.1, -29.8, -21.0, -28.7),
                           p1 = c(4.71, 6.36, 4.62, 6.23),
                           a2 = c(-26.7, -31.0, -25.7, -30.0),
                           p2 = c(5.64, 6.62, 5.47, 6.42))
head(risk_table_A)
```

### 1. Calculate the underlying risks for coronary heart disease and non-coronary cardiovascular disease

Calculate the underlying risks for coronary heart disease and non-coronary cardiovascular disease separately for the person's age now and for their age in ten years time, using values from $\alpha$ and $p$ shown in table A. The underlying survival probability is denoted as $S_0$ is calculated by

\begin{align}
S_0(age) = exp\{-(exp(a))(age - 20)^p\}\\
S_0(age+10) = exp\{-(exp(a))(age - 10)^p\}
\end{align}

```{r}
survival_probability <- function(age, a, p){
  #' Calculates the survivability probability using coefficients a and p
  #' 
  #' @param age The age of the patient 
  #' @param a Coefficient a
  #' @param p Coefficient p 
  S0_age <- exp(-(exp(a)) * ((age-20)**p))
  S0_age_10 <- exp(-(exp(a)) * ((age-10)**p))
  result <- list(S0_age, S0_age_10)
  return(result)
}
```


### 2. Calculate weighted sum $w$ of risk factors

Using coefficients in table B, calculate the weighted sum, $w$, of the risk factors cholestrol, smoking and systolic blood pressure. Two weighted sums, one for coronary heart disease and one for non-coronary heart disease, will be calculated. 


```{r}
# Smoker, Chol, SBP
weighted_sum_risk_factor <- function(smoker, cholestrol, SBP, weights){
  w <- (weights[1] * smoker) + 
       (weights[2] * (cholestrol - 6)) + 
       (weights[3] * (SBP-120))
    
  return(w)
}
```

### 3. Combining underlying risk
```{r}
combined_risk <- function(weight, risks){
  S_age = risks[[1]] ** exp(weight)
  S_age_10 = risks[[2]] ** exp(weight)
  S10_age = S_age_10 / S_age
  Risk_10 = 1 - S10_age
  return(Risk_10)
}
```

### Test Input 
Patient: 
* Age 55
* Low risk
* Men
* SBP 120
* Cholestrol 6
* Smoker 1

```{r}
CHD_result = survival_probability(55, -22.1, 4.71)
CHD_result
```

```{r}
NON_CHD_result = survival_probability(55, -26.7, 5.64)
NON_CHD_result
```

```{r}
# Smoker, Chol, SBP
weights = c(0.71, 0.24, 0.018)
CHD_weight = weighted_sum_risk_factor(1, 4, 120, weights)
CHD_weight
```

```{r}
weights = c(0.63, 0.02, 0.022)
NON_CHD_weight = weighted_sum_risk_factor(1, 4, 120, weights)
NON_CHD_weight
```

Combine underlying risk

```{r}
CHD_Risk_10 = combined_risk(CHD_weight, CHD_result)
CHD_Risk_10
```

```{r}
NON_CHD_Risk_10 = combined_risk(NON_CHD_weight, NON_CHD_result)
NON_CHD_Risk_10
```

```{r}
total_risk = CHD_Risk_10 + NON_CHD_Risk_10
total_risk
```